---
title: "Northeast enrollment modeling"
author: "John Baker"
format:
  html:
    code-overflow: wrap
    embed-resources: true
---

```{r}
rm(list = ls())
```

```{r, packages}
library(tidyverse)
library(tidymodels)
library(leaps)
library(car)
library(doParallel)
```

adding this to speed up code (from AI, code was taking forever)

```{r, speed up code}
cores = max(1, parallel::detectCores() - 1)
cl = parallel::makePSOCKcluster(cores)
registerDoParallel(cl)
```

```{r, read info}
ne = read_csv("northeast_census_over_time_modeling_gender.csv",
col_types = cols(ZIP = col_character(), .default = col_guess())) |>
    mutate(state = as.factor(state))
```

```{r, train test}
set.seed(88)

train_data = ne |> filter(year < 2024)
test_data = ne |> filter(year == 2024)
```

```{r, validate grouped by ZIP}
set.seed(88)

folds_grouped = group_vfold_cv(train_data, group = ZIP, v = 10)
```

Before doing recipes, let's do some subset work

```{r, ideal reduced model}
subset_data = train_data |>
  select(
    private_k12,
    # Economic characteristics
    median_hh_income_inflation_adjusted,
    median_home_value_inflation_adjusted,
    # Adult educational attainment
    pct_adults_hs_plus,
    pct_adults_bachelors_plus,
    pct_adults_grad_plus,
    # Household characteristics
    total_households,
    pct_households_with_children,
    # Child population by age
    age_0_4,
    age_5_9,
    age_10_14,
    age_15_17,
    # Gender composition
    pct_school_age_male)

subset_fit = regsubsets(
  private_k12 ~ .,
  data = subset_data,
  nvmax = 12,
  method = "exhaustive")

subset_summary = summary(subset_fit)
subset_comparison = tibble(
  predictors = 1:12,
  adjusted_r2 = subset_summary$adjr2,
  cp = subset_summary$cp,
  bic = subset_summary$bic)

subset_comparison
```

Which works best?

```{r}
best_adjr2 = which.max(subset_summary$adjr2)
best_cp = which.min(subset_summary$cp)
best_bic = which.min(subset_summary$bic)

best_adjr2
best_cp
best_bic

coef(subset_fit, id = best_adjr2)

coef(subset_fit, id = best_cp)

coef(subset_fit, id = best_bic)
```

Looks like 11! Will try three models to be sure.

```{r, recipes for OLS}

# Recipe A: All features

recipe_a = recipe(
  private_k12 ~ median_hh_income_inflation_adjusted + median_home_value_inflation_adjusted +
    school_age_population + pct_adults_hs_plus + pct_adults_bachelors_plus +
    pct_adults_grad_plus + total_households + pct_households_with_children + age_0_4 +
    state,
  data = train_data) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())

# Recipe B: 11 numeric predictors + state, no gender

recipe_b = recipe(
  private_k12 ~ median_hh_income_inflation_adjusted + median_home_value_inflation_adjusted +
    pct_adults_hs_plus + pct_adults_bachelors_plus + pct_adults_grad_plus + total_households +
    pct_households_with_children + age_0_4 + age_5_9 + age_10_14 + age_15_17 + state,
  data = train_data) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())

# Recipe C: Same as B, no bachelors for parents

recipe_c = recipe(
  private_k12 ~ median_hh_income_inflation_adjusted + median_home_value_inflation_adjusted +
    pct_adults_hs_plus + pct_adults_grad_plus + total_households +
    pct_households_with_children + age_0_4 + age_5_9 + age_10_14 + age_15_17 + state,
  data = train_data) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())
```

```{r, prep and metrics}

missing_a = prep(recipe_a) |>
  bake(new_data = NULL) |>
  summarize(across(everything(), ~sum(is.na(.x))))

missing_b = prep(recipe_b) |>
  bake(new_data = NULL) |>
  summarize(across(everything(), ~sum(is.na(.x))))

missing_c = prep(recipe_c) |>
  bake(new_data = NULL) |>
  summarize(across(everything(), ~sum(is.na(.x))))

missing_a
missing_b
missing_c

reg_metrics = metric_set(rmse, rsq, mae)
reg_metrics
```

```{r, workflows}
set.seed(88)
lr_spec = linear_reg(mode = "regression") |> set_engine("lm")

lr_wf_a = workflow() |> add_recipe(recipe_a) |> add_model(lr_spec)
lr_wf_b = workflow() |> add_recipe(recipe_b) |> add_model(lr_spec)
lr_wf_c = workflow() |> add_recipe(recipe_c) |> add_model(lr_spec)

lr_cv_a = fit_resamples(lr_wf_a,
  resamples = folds_grouped,
  metrics = reg_metrics,
  control = control_resamples(save_pred = TRUE))
lr_cv_b = fit_resamples(lr_wf_b,
  resamples = folds_grouped,
  metrics = reg_metrics,
  control = control_resamples(save_pred = TRUE))
lr_cv_c = fit_resamples(lr_wf_c,
  resamples = folds_grouped,
  metrics = reg_metrics,
  control = control_resamples(save_pred = TRUE))
```

```{r, compare OLS metrics}
linreg_metrics = bind_rows(
  collect_metrics(lr_cv_a) |> mutate(model = "OLS A - Aggregate ages"),
  collect_metrics(lr_cv_b) |> mutate(model = "OLS B - Full individual ages"),
  collect_metrics(lr_cv_c) |> mutate(model = "OLS C - Reduced individual ages")) |>
  dplyr::select(model, .metric, mean, std_err) |>
  arrange(.metric, mean)

linreg_metrics
```

```{r, fit to training data}
ols_fit_a = lr_wf_a |> fit(data = train_data)
ols_fit_b = lr_wf_b |> fit(data = train_data)
ols_fit_c = lr_wf_c |> fit(data = train_data)
ols_fit_a |> extract_fit_engine() |> summary()
ols_fit_b |> extract_fit_engine() |> summary()
ols_fit_c |> extract_fit_engine() |> summary()
```

I'm curious about outliers.

```{r, outliers}
extreme_five = c(
  "08701",  # Lakewood, NJ
  "11219",  # Brooklyn, NY
  "10977",  # Spring Valley, NY
  "10950",  # Monroe, NY
  "10952")   # Monsey, NY

midatlantic_high = test_data |>
  filter(state %in% c("DC", "MD", "VA")) |>
  arrange(desc(private_k12)) |>
  group_by(state) |>
  slice_head(n = 3) |>
  ungroup() |>
  dplyr::select(ZIP, zip_place_name, state, private_k12, school_age_population)

midatlantic_high

diagnostic_zips = c(extreme_five, midatlantic_high$ZIP) |> unique()
train_lookup = train_data |>
  mutate(.row = row_number()) |>
  dplyr::select(.row, ZIP, year, state, zip_place_name)

pred_a = collect_predictions(lr_cv_a) |>
  left_join(train_lookup, by = ".row") |>
  mutate(model = "A - Aggregate ages")
pred_b = collect_predictions(lr_cv_b) |>
  left_join(train_lookup, by = ".row") |>
  mutate(model = "B - Full individual ages")
pred_c = collect_predictions(lr_cv_c) |>
  left_join(train_lookup, by = ".row") |>
  mutate(model = "C - Reduced individual ages")

all_ols_predictions = bind_rows(pred_a, pred_b, pred_c)
all_ols_predictions = all_ols_predictions |>
  mutate(market_group = case_when(
    ZIP %in% extreme_five ~ "Original five extreme ZIPs",
    ZIP %in% diagnostic_zips ~ "High DC / MD / VA ZIPs",
    TRUE ~ "Other ZIPs"))

extreme_comparison = all_ols_predictions |>
  group_by(model, market_group) |>
  summarize(
    n = n(),
    mae = mae_vec(truth = private_k12, estimate = .pred),
    rmse = rmse_vec(truth = private_k12, estimate = .pred),
    mean_error = mean(.pred - private_k12),
    .groups = "drop")

extreme_comparison
```

```{r}
zip_comparison = all_ols_predictions |>
  filter(ZIP %in% diagnostic_zips) |>
  group_by(model, ZIP, zip_place_name, state) |>
  summarize(
    actual_mean = mean(private_k12),
    predicted_mean = mean(.pred),
    mean_error = mean(.pred - private_k12),
    mae = mean(abs(.pred - private_k12)),
    .groups = "drop") |>
  arrange(ZIP, model)

zip_comparison
```

So I have some concerns about multicollinearity here. Let's check!

```{r}
vif_a = ols_fit_a |>
  extract_fit_engine() |>
  car::vif()

vif_c = ols_fit_c |>
  extract_fit_engine() |>
  car::vif()

# Compare VIF values
tibble(
  variable = names(vif_a),
  VIF = as.numeric(vif_a)) |>
  arrange(desc(VIF))

tibble(
  variable = names(vif_c),
  VIF = as.numeric(vif_c)) |>
  arrange(desc(VIF))
```

Let's try an additional Model D that could improve on these high VIF values

```{r}
set.seed(88)

recipe_d = recipe(
  private_k12 ~ median_hh_income_inflation_adjusted + median_home_value_inflation_adjusted +
    school_age_population + pct_adults_hs_plus + pct_adults_grad_plus +
    total_households + pct_households_with_children + state,
  data = train_data) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())

lr_wf_d = workflow() |>
  add_recipe(recipe_d) |>
  add_model(lr_spec)

lr_cv_d = fit_resamples(lr_wf_d,
  resamples = folds_grouped,
  metrics = reg_metrics,
  control = control_resamples(save_pred = TRUE))

ols_fit_d = lr_wf_d |>
  fit(data = train_data)

vif_d = ols_fit_d |>
  extract_fit_engine() |>
  car::vif()

tibble(
  variable = names(vif_d),
  VIF = as.numeric(vif_d)) |>
  arrange(desc(VIF))
```

```{r}
linreg_metrics_acd = bind_rows(
  collect_metrics(lr_cv_a) |>
    mutate(model = "OLS A - Aggregate ages"),

  collect_metrics(lr_cv_c) |>
    mutate(model = "OLS C - Reduced individual ages"),

  collect_metrics(lr_cv_d) |>
    mutate(model = "OLS D - Reduced aggregate ages")
) |>
  select(
    model,
    .metric,
    mean,
    std_err
  ) |>
  arrange(.metric, mean)

linreg_metrics_acd
```

Model D improves upon model A in MAE, but worse in RMSE. I'm curious to use A, C, and D in a regularized regression. We see outliers, especially in towns like Lakewood Township, NJ where a lot of Orthodox Jews live who are far less likely to attend public schools. Model D does not perform as well on the outliers but performs better on most ZIPs. This is more interesting to me, I'd rather it predict 99% right and be off on the outliers that don't follow other patterns.


```{r, regularized setup}
recipe_a_reg = recipe_a |>
  step_normalize(all_numeric_predictors())
recipe_c_reg = recipe_c |>
  step_normalize(all_numeric_predictors())
recipe_d_reg = recipe_d |>
  step_normalize(all_numeric_predictors())

reg_spec = linear_reg(mode = "regression", penalty = tune(), mixture = tune()) |>
  set_engine("glmnet")

# workflows

reg_wf_a = workflow() |> add_recipe(recipe_a_reg) |> add_model(reg_spec)
reg_wf_c = workflow() |> add_recipe(recipe_c_reg) |> add_model(reg_spec)
reg_wf_d = workflow() |> add_recipe(recipe_d_reg) |> add_model(reg_spec)

# tuning grid

reg_grid = crossing(
  penalty = 10^seq(-4, 3, length.out = 20),
  mixture = c(0, 0.5, 1))
reg_grid
```


```{r, tuning regularized models}
set.seed(88)

reg_cv_a = tune_grid(
  reg_wf_a,
  resamples = folds_grouped,
  grid = reg_grid,
  metrics = reg_metrics,
  control = control_grid())

reg_cv_c = tune_grid(
  reg_wf_c,
  resamples = folds_grouped,
  grid = reg_grid,
  metrics = reg_metrics,
  control = control_grid())

reg_cv_d = tune_grid(
  reg_wf_d,
  resamples = folds_grouped,
  grid = reg_grid,
  metrics = reg_metrics,
  control = control_grid())
```


```{r, regularization results}
autoplot(reg_cv_a)
autoplot(reg_cv_c)
autoplot(reg_cv_d)

# best metrics, MAE and RMSE

show_best(reg_cv_a, metric = "mae", n = 5)
show_best(reg_cv_c, metric = "mae", n = 5)
show_best(reg_cv_d, metric = "mae", n = 5)
show_best(reg_cv_a, metric = "rmse", n = 5)
show_best(reg_cv_c, metric = "rmse", n = 5)
show_best(reg_cv_d, metric = "rmse", n = 5)

```


```{r}

# Finalize the MAE-selected configurations

best_a = select_best(reg_cv_a, metric = "mae")
best_c = select_best(reg_cv_c, metric = "mae")
best_d = select_best(reg_cv_d, metric = "mae")

reg_fit_a = reg_wf_a |>
  finalize_workflow(best_a) |>
  fit(data = train_data)
reg_fit_c = reg_wf_c |>
  finalize_workflow(best_c) |>
  fit(data = train_data)
reg_fit_d = reg_wf_d |>
  finalize_workflow(best_d) |>
  fit(data = train_data)

# Examine variation in predictions

bind_rows(
  predict(reg_fit_a, new_data = train_data) |>
    mutate(model = "A"),
  predict(reg_fit_c, new_data = train_data) |>
    mutate(model = "C"),
  predict(reg_fit_d, new_data = train_data) |>
    mutate(model = "D")) |>
  group_by(model) |>
  summarize(
    min_prediction = min(.pred),
    median_prediction = median(.pred),
    max_prediction = max(.pred),
    prediction_sd = sd(.pred),
    .groups = "drop"
  )
```

```{r}
# Training predictions for both models

reg_predictions = bind_rows(
  train_data |>
    select(ZIP, year, private_k12) |>
    bind_cols(predict(reg_fit_a, new_data = train_data)) |>
    mutate(model = "A"),
  train_data |>
    select(ZIP, year, private_k12) |>
    bind_cols(predict(reg_fit_c, new_data = train_data)) |>
    mutate(model = "C"),
   train_data |>
    select(ZIP, year, private_k12) |>
    bind_cols(predict(reg_fit_d, new_data = train_data)) |>
    mutate(model = "D"), )

# Count impossible predictions

reg_predictions |>
  group_by(model) |>
  summarize(
    negative_predictions = sum(.pred < 0),
    pct_negative = mean(.pred < 0) * 100,
    .groups = "drop")
```

```{r}
reg_metrics_a = collect_metrics(reg_cv_a) |>
  semi_join(best_a, by = c("penalty", "mixture")) |>
  mutate(model = "Regularized A")
reg_metrics_c = collect_metrics(reg_cv_c) |>
  semi_join(best_c, by = c("penalty", "mixture")) |>
  mutate(model = "Regularized C")
reg_metrics_d = collect_metrics(reg_cv_d) |>
  semi_join(best_d, by = c("penalty", "mixture")) |>
  mutate(model = "Regularized D")

ols_vs_regularized = bind_rows(
  collect_metrics(lr_cv_a) |>
    mutate(model = "OLS A"),
  collect_metrics(lr_cv_c) |>
    mutate(model = "OLS C"),
  collect_metrics(lr_cv_d) |>
    mutate(model = "OLS D"),
  reg_metrics_a,
  reg_metrics_c,
  reg_metrics_d) |>
  dplyr::select(model, .metric, mean, std_err) |>
  arrange(.metric, mean)

ols_vs_regularized
```

```{r, coef check}
best_a
best_c
best_d

tidy(reg_fit_a) |>
  filter(term != "(Intercept)") |>
  arrange(desc(abs(estimate)))

tidy(reg_fit_c) |>
  filter(term != "(Intercept)") |>
  arrange(desc(abs(estimate)))

tidy(reg_fit_d) |>
  filter(term != "(Intercept)") |>
  arrange(desc(abs(estimate)))
```

I like, based on MAE and RSQ, Regularized model C most. Interestingly, RMSE and MAE give different best models.

### Ensemble Methods

Let's move on and do our ensemble methods with trees, random forests, and boosting.

```{r, boosting preprocessing}
tune_xgboost_spec = boost_tree(
  mode = "regression",
  trees = tune(),
  learn_rate = tune(),
  tree_depth = tune()) |>
  set_engine("xgboost")

tune_xgboost_wf_a = workflow() |>
  add_recipe(recipe_a) |>
  add_model(tune_xgboost_spec)
tune_xgboost_wf_c = workflow() |>
  add_recipe(recipe_c) |>
  add_model(tune_xgboost_spec)
tune_xgboost_wf_d = workflow() |>
  add_recipe(recipe_d) |>
  add_model(tune_xgboost_spec)

xgboost_params_a = extract_parameter_set_dials(tune_xgboost_wf_a) |>
  update(trees = trees(c(50, 500)),
    learn_rate = learn_rate(c(-3, -1)),
    tree_depth = tree_depth(c(1, 5)))
xgboost_params_c = extract_parameter_set_dials(tune_xgboost_wf_c) |>
  update(trees = trees(c(50, 500)),
    learn_rate = learn_rate(c(-3, -1)),
    tree_depth = tree_depth(c(1, 5)))
xgboost_params_d = extract_parameter_set_dials(tune_xgboost_wf_d) |>
  update(trees = trees(c(50, 500)),
    learn_rate = learn_rate(c(-3, -1)),
    tree_depth = tree_depth(c(1, 5)))


```

```{r, boosting tuning}
tune_results_xgboost_a = tune_grid(
  tune_xgboost_wf_a,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(xgboost_params_a, size = 20))
tune_results_xgboost_c = tune_grid(
  tune_xgboost_wf_c,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(xgboost_params_a, size = 20))
tune_results_xgboost_d = tune_grid(
  tune_xgboost_wf_d,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(xgboost_params_a, size = 20))
```

```{r, boosting metrics}
autoplot(tune_results_xgboost_a)
autoplot(tune_results_xgboost_c)
autoplot(tune_results_xgboost_d)

show_best(tune_results_xgboost_a, metric = "mae", n = 5)
show_best(tune_results_xgboost_a, metric = "rmse", n = 5)
show_best(tune_results_xgboost_a, metric = "rsq", n = 5)
show_best(tune_results_xgboost_c, metric = "mae", n = 5)
show_best(tune_results_xgboost_c, metric = "rmse", n = 5)
show_best(tune_results_xgboost_c, metric = "rsq", n = 5)
show_best(tune_results_xgboost_d, metric = "mae", n = 5)
show_best(tune_results_xgboost_d, metric = "rmse", n = 5)
show_best(tune_results_xgboost_d, metric = "rsq", n = 5)

best_xgboost_a = select_best(tune_results_xgboost_a, metric = "mae")
best_xgboost_c = select_best(tune_results_xgboost_c, metric = "mae")
best_xgboost_d = select_best(tune_results_xgboost_d, metric = "mae")

best_xgboost_a
best_xgboost_c
best_xgboost_d
```

We can now move onto random forests

```{r, rf preprocessing}
tune_rf_spec = rand_forest(
  mode = "regression",
  mtry = tune(),
  min_n = tune(),
  trees = 300) |>
  set_engine( "ranger", importance = "impurity", num.threads = 1)

tune_rf_wf_a = workflow() |>
  add_recipe(recipe_a) |>
  add_model(tune_rf_spec)
tune_rf_wf_c = workflow() |>
  add_recipe(recipe_c) |>
  add_model(tune_rf_spec)
tune_rf_wf_d = workflow() |>
  add_recipe(recipe_d) |>
  add_model(tune_rf_spec)

processed_train_a = prep(recipe_a) |>
  bake(new_data = NULL) |>
  select(-private_k12)
processed_train_c = prep(recipe_c) |>
  bake(new_data = NULL) |>
  select(-private_k12)
processed_train_d = prep(recipe_d) |>
  bake(new_data = NULL) |>
  select(-private_k12)

rf_params_a = extract_parameter_set_dials(tune_rf_wf_a) |>
  finalize(processed_train_a) |>
  update(min_n = min_n(c(2, 25)))
rf_params_c = extract_parameter_set_dials(tune_rf_wf_c) |>
  finalize(processed_train_c) |>
  update(min_n = min_n(c(2, 25)))
rf_params_d = extract_parameter_set_dials(tune_rf_wf_d) |>
  finalize(processed_train_d) |>
  update(min_n = min_n(c(2, 25)))
```

```{r, rf tuning}
set.seed(88)

tune_results_rf_a = tune_grid(
  tune_rf_wf_a,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(rf_params_a, size = 20))


```

```{r, tune rd c}
set.seed(88)

tune_results_rf_c = tune_grid(
  tune_rf_wf_c,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(rf_params_c, size = 20))
```

```{r, tun rf d}
set.seed(88)

tune_results_rf_d = tune_grid(
  tune_rf_wf_d,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(rf_params_d, size = 20))
```

```{r}
autoplot(tune_results_rf_a)
autoplot(tune_results_rf_c)
autoplot(tune_results_rf_d)

show_best(tune_results_rf_a, metric = "mae", n = 5)
show_best(tune_results_rf_a, metric = "rmse", n = 5)
show_best(tune_results_rf_a, metric = "rsq", n = 5)
show_best(tune_results_rf_c, metric = "mae", n = 5)
show_best(tune_results_rf_c, metric = "rmse", n = 5)
show_best(tune_results_rf_c, metric = "rsq", n = 5)
show_best(tune_results_rf_d, metric = "mae", n = 5)
show_best(tune_results_rf_d, metric = "rmse", n = 5)
show_best(tune_results_rf_d, metric = "rsq", n = 5)
```

Random Forests seem really strong!!! That's good. Let's investigate some more.

```{r}
best_rf_a = select_best(tune_results_rf_a, metric = "mae")
best_rf_c = select_best(tune_results_rf_c, metric = "mae")
best_rf_d = select_best(tune_results_rf_d, metric = "mae")

collect_metrics(tune_results_rf_a) |>
  semi_join(best_rf_a, by = names(best_rf_a)) |>
  mutate(model = "RF A")

rf_best_metrics = bind_rows(
  collect_metrics(tune_results_rf_a) |>
    semi_join(best_rf_a, by = names(best_rf_a)) |>
   mutate(model = "RF A"),
  collect_metrics(tune_results_rf_c) |>
    semi_join(best_rf_c, by = names(best_rf_c)) |>
    mutate(model = "RF C"),
  collect_metrics(tune_results_rf_d) |>
    semi_join(best_rf_d, by = names(best_rf_d)) |>
    mutate(model = "RF D"))

rf_best_metrics
```

Let's try SVM now (at least a radial version of it, already have a linear/non-linear, lets get radial)

```{r}
set.seed(88)

svm_spec = svm_rbf(
  mode = "regression",
  cost = tune(),
  rbf_sigma = tune()) |>
  set_engine("kernlab")
svm_wf_a = workflow() |>
  add_recipe(recipe_a_reg) |>
  add_model(svm_spec)
svm_params_a = extract_parameter_set_dials(svm_wf_a)

tune_results_svm_a = tune_grid(
  svm_wf_a,
  resamples = folds_grouped,
  metrics = reg_metrics,
  grid = grid_random(svm_params_a, size = 10))
```

```{r}
show_best(tune_results_svm_a, metric = "mae", n = 5)

best_svm_a = select_best(
  tune_results_svm_a,
  metric = "mae")

collect_metrics(tune_results_svm_a) |>
  semi_join(
    best_svm_a,
    by = names(best_svm_a))

final_svm_wf = svm_wf_a |>
  finalize_workflow(best_svm_a)

final_svm_fit = final_svm_wf |>
  fit(data = train_data)
```

### Predict on testing data

```{r}
final_linreg_a = lr_wf_a |>
  fit(data = train_data)
final_xgboost_a = tune_xgboost_wf_a |>
  finalize_workflow(best_xgboost_a) |>
  fit(data = train_data)
final_rf_a = tune_rf_wf_a |>
  finalize_workflow(best_rf_a) |>
  fit(data = train_data)
final_svm_a = svm_wf_a |>
  finalize_workflow(best_svm_a) |>
  fit(data = train_data)

linreg_test = final_linreg_a |>
  augment(new_data = test_data)
xgboost_test = final_xgboost_a |>
  augment(new_data = test_data)
rf_test = final_rf_a |>
  augment(new_data = test_data)
svm_test = final_svm_a |>
  augment(new_data = test_data)
```

Now we calculate our metrics

```{r}
test_results = bind_rows(
  linreg_test |> 
    reg_metrics(truth = private_k12, estimate = .pred) |>
    mutate(model = "Linear Regression"),
  xgboost_test |>
    reg_metrics(truth = private_k12, estimate = .pred) |>
    mutate(model = "XGBoost"),
  rf_test |>
    reg_metrics(truth = private_k12, estimate = .pred) |>
    mutate(model = "Random Forest"),
  svm_test |>
    reg_metrics(truth = private_k12, estimate = .pred) |>
    mutate(model = "RBF SVM"))

test_results
```

```{r}
test_results_table = test_results |>
  dplyr::select(model, .metric, .estimate) |>
  pivot_wider(names_from = .metric, values_from = .estimate) |>
  arrange(mae)

test_results_table
```

```{r}
rf_test_errors = rf_test |>
  mutate(
    error = .pred - private_k12,
    abs_error = abs(error))

rf_test_errors |>
  summarize(
    mean_absolute_error = mean(abs_error),
    median_absolute_error = median(abs_error),
    q25_absolute_error = quantile(abs_error, 0.25),
    q75_absolute_error = quantile(abs_error, 0.75),
    q90_absolute_error = quantile(abs_error, 0.90))

rf_test_errors = rf_test_errors |>
  mutate(
    enrollment_group = case_when(
      private_k12 < 50 ~ "Under 50",
      private_k12 < 100 ~ "50-99",
      private_k12 < 250 ~ "100-249",
      private_k12 < 500 ~ "250-499",
      private_k12 < 1000 ~ "500-999",
      TRUE ~ "1000+"))

rf_test_errors |>
  group_by(enrollment_group) |>
  summarize(
    n = n(),
    mean_actual = mean(private_k12),
    mae = mean(abs_error),
    median_abs_error = median(abs_error),
    rmse = sqrt(mean(error^2))
  )
```

This looks really good once zips get big but not as good for smaller ones.
It looks like there are likely some predicted negative values, which can't be possible in real life.

```{r}
rf_test |> summarize(
  negative_predictions = sum(.pred < 0),
  pct_negative = mean(.pred < 0) * 100,
  minimum_prediction = min(.pred))

rf_test
```


```{r}

set.seed(88)

train_data_rate = train_data |>
  mutate(private_rate = private_k12 / school_age_population)

test_data_rate = test_data |>
  mutate(private_rate = private_k12 / school_age_population)

folds_grouped_rate = group_vfold_cv(
  train_data_rate, group = ZIP, v = 10)

recipe_a_rate = recipe(
  private_rate ~
    median_hh_income_inflation_adjusted + median_home_value_inflation_adjusted +
    school_age_population + pct_adults_hs_plus + pct_adults_bachelors_plus +
    pct_adults_grad_plus + total_households + pct_households_with_children +
    age_0_4 + state,
  data = train_data_rate) |>
  step_dummy(all_nominal_predictors()) |>
  step_zv(all_predictors())

rf_rate_spec = rand_forest(
  mode = "regression",
  mtry = tune(),
  min_n = tune(),
  trees = 300) |>
  set_engine("ranger", importance = "impurity")

rf_rate_wf = workflow() |>
  add_recipe(recipe_a_rate) |>
  add_model(rf_rate_spec)
rf_rate_prepped = prep(recipe_a_rate)
rf_rate_num_predictors = rf_rate_prepped |>
  bake(new_data = NULL) |>
  select(-private_rate) |>
  ncol()
rf_rate_params = extract_parameter_set_dials(rf_rate_wf) |>
  update(
    mtry = mtry(c(1, rf_rate_num_predictors)),
    min_n = min_n(c(2, 25)))

```

Now we tune based on this new data

```{r}
set.seed(88)

tune_results_rf_rate = tune_grid(
  rf_rate_wf,
  resamples = folds_grouped_rate,
  metrics = reg_metrics,
  grid = grid_random(
    rf_rate_params,
    size = 20))

show_best(tune_results_rf_rate, metric = "mae", n = 5)
best_rf_rate = select_best(tune_results_rf_rate, metric = "mae")
best_rf_rate

collect_metrics(tune_results_rf_rate) |>
  semi_join(
    best_rf_rate,
    by = names(best_rf_rate))

final_rf_rate_wf = rf_rate_wf |>
  finalize_workflow(best_rf_rate)

final_rf_rate = final_rf_rate_wf |>
  fit(data = train_data_rate)
```

Now we can predict on 2024 data

```{r}
rf_rate_test = final_rf_rate |>
  augment(new_data = test_data_rate) |>
  mutate(.pred_count = .pred * school_age_population)
rf_rate_test_metrics = rf_rate_test |>
  reg_metrics(truth = private_k12, estimate = .pred_count)
rf_rate_test_metrics

rf_rate_comparison = bind_rows(
  rf_test |>
    reg_metrics(truth = private_k12, estimate = .pred) |>
    mutate(model = "RF Raw Count"),
  rf_rate_test |>
    reg_metrics(truth = private_k12, estimate = .pred_count) |>
    mutate(model = "RF Rate")) |>
    dplyr::select(model, .metric, .estimate) |>
    pivot_wider(
      names_from = .metric,
      values_from = .estimate)

rf_rate_comparison
```

This should really help our tails, as the outliers from the previous model were really off.

```{r}
rf_rate_errors = rf_rate_test |>
  mutate(
    error = .pred_count - private_k12,
    abs_error = abs(error))

rf_rate_errors |>
  summarize(
    mean_absolute_error = mean(abs_error),
    median_absolute_error = median(abs_error),
    q25_absolute_error = quantile(abs_error, 0.25),
    q75_absolute_error = quantile(abs_error, 0.75),
    q90_absolute_error = quantile(abs_error, 0.90))
```

It does better with the outliers, hooray! The Random Forest capturing rate of private school enrollment rather than raw count is strongest.

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

Stop our clustering!

```{r}
parallel::stopCluster(cl)
```

```{r}
dataframe = read_csv("northeast_census_data_gender.csv")
```

```{r}
ne = ne |>
  mutate(actual_private_rate = private_k12 / school_age_population)

rf_rate_features = rf_rate_test |>
  transmute(ZIP, year, projected_private_rate = .pred, projected_private_k12 = .pred_count)

ne = ne |>
  left_join(rf_rate_features, by = c("ZIP", "year"))
```

```{r}
write_csv(ne, 'northeast_data_modeling_final.csv')
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```